---
title: "codec"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{codec}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

### Dependence and Conditional Dependence
For random variable $Y$ and random vectors $Z$ and $X$, $T(Y, Z\mid X)\in[0, 1]$, conditional dependence coefficient, gives us a measure of dependence of $Y$ and $Z$ given $X$, which is zero if and only if $Y$ is independent of $Z$ given $X$ and is 1 if and only if $Y$ is a function of $Z$ given $X$. Function `codec` estimates this value. The default value for $X$ is `NULL` and if is not provided by the user, it gives the dependence measure of $Y$ on $Z$. For more details on the definition of $T$ and its properties, see the paper [*A Simple Measure Of Conditional Dependence*](https://arxiv.org/pdf/1910.12327.pdf).

Below you can see a simple example of this measure.
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(fig.width=7, fig.height=5) 
```

```{r setup, results='hide', message=FALSE, warning=FALSE}
library(FOCI)
```

```{r}
n = 10000
x1 = matrix(runif(n), ncol = 1)
x2 = matrix(runif(n), ncol = 1)
x3 = matrix(runif(n), ncol = 1)
y = (x1 + x2 + x3) %% 1
# y is independent of each of x1 and x2 and x3 
codec(y, x1)
codec(y, x2)
codec(y, x3)

# y is independent of the pair (x1, x2)
codec(y, cbind(x1, x2))

# y is a function of (x1, x2, x3)
codec(y, cbind(x1, x2, x3))

# conditional on x3, y is a function of (x1, x2)
codec(y, cbind(x1, x2), x3)
# conditional on x3, y is independent of x1
codec(y, x1, x3)

```

